### 2018-04-10 latex/icpp18/algorithm.tex
reintroduced figure1 and remove k=16 references in experiments

%\begin{figure}
%\caption{A 5x5 checkerboard partition of a sparse matrix.}
%\label{fig:partition}
%\includegraphics[width=0.8\linewidth]{figures/matrix-partition.png}
%\end{figure}

@@

Traditional checkerboard hypergraph partitioning aims to partition the matrix $\Am$ into $P_r$ row slices first, and $P_c$ column slices next to obtain an $P_r \times P_c$ checkerboard partition.
The first partitioning phase is done using a column-net hypergraph model, in which for each row $\Am(i, :)$, a vertex $v_i$ with weight equaling to the number of nonzeros in $\Am(i, :)$ is created.
Each column $j$ is represented with a hyperedge $h_j$ and for each nonzero $(i, j) \in \Am$, which implies a dependency to $\Hm(:, j)$ in computing $\Wmt(i, :)$ at \cref{line:ah}, we connect $v_i$ to $h_j$.
This hypergraph is partitioned into $P_r$ parts giving the row partition of the checkerboard topology.
The second partitioning phase uses a row-net hypergraph model induced by this row partition, where each column is represented with a vertex with $P_r$ weights corresponding to the number of nonzeros in that column in all $P_r$ row segments.
Partitioning this hypergraph into $P_c$ parts finalizes the $P_r \times P_c$ checkerboard partitionin by balancing the weights~(number of nonzeros of $\Am_{(r, c)}$) of each part while minimizing the communication volume.
In \textbf{CH2},

>>>>

\begin{figure}
\caption{A 5x5 checkerboard partition of a sparse matrix.}
\label{fig:partition}
\includegraphics[width=0.8\linewidth]{figures/matrix-partition.png}
\end{figure}

@@

Traditional checkerboard hypergraph partitioning aims to partition the matrix $\Am$ into $P_r$ row slices first, and $P_c$ column slices next to obtain an $P_r \times P_c$ checkerboard partition.
The first partitioning phase is done using a column-net hypergraph model, in which for each row $\Am(i, :)$, a vertex $v_i$ with weight equaling to the number of nonzeros in $\Am(i, :)$ is created.
Each column $j$ is represented with a hyperedge $h_j$ and for each nonzero $(i, j) \in \Am$, which implies a dependency to $\Hm(:, j)$ in computing $\Wmt(i, :)$ at \cref{line:ah}, we connect $v_i$ to $h_j$.
This hypergraph is partitioned into $P_r$ parts giving the row partition of the checkerboard topology.
The second partitioning phase uses a row-net hypergraph model induced by this row partition, where each column is represented with a vertex with $P_r$ weights corresponding to the number of nonzeros in that column in all $P_r$ row segments.
Partitioning this hypergraph into $P_c$ parts finalizes the $P_r \times P_c$ checkerboard partitionin by balancing the weights~(number of nonzeros of $\Am_{(r, c)}$) of each part while minimizing the communication volume.
In the context of NMF, one potential issue arises when the matrix has a large variance in the numbe r of nonzeros in its rows/columns, which in turns yields unbalanced row/column strides.
This in turn creates an imbalance in the NLS computations as rows/cols of $\HH$/$\WW$ are distributed to the processes in the same stride.
To alleviate this issue, we modify this scheme slightly, which we call \textbf{CH2}.
In both row and column partitioning phases, we add an additional constant weight to vertices.
Balancing this additional constraint in hypergraph partitioning is expected to prevent such imbalanced strides.
This partitioning model succesfully grasps the computation~(both SpMM and NLS) and communication requirements using checkerboard topology for sparse NMF, yet it is very costly to compute in practice due to high number of constraints in the row-net hypergraph.
